---
title: "P8105_HW1"
author: "Ngoc Duong_nqd2000"
date: "9/20/2019"
output: github_document
---

## Problem 1
```{r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)

library(tidyverse)
library(dplyr)
library(knitr)
library(base)

#set seed for random sampling
set.seed(77)

#create tibble dataframe containing a sample of 8 following a normal distribution with mean = 0 and sd = 1 
df <- tibble(random_sample = rnorm(8,mean = 0, sd =1),
            #logical vector for if each element is greater than 0 
            gr_th_0 = random_sample > 0, 
            #make a character vector of length 8
            num_char = as.character(c(1,2,3,4,5,6,7,8)), 
            #make a factor vector of length 8
            level = factor(c("Low", "Medium","High","Low","Medium","High","Medium","High")) 
            )

#Taking the mean of all vectors in tibble dataframe df
mean(pull(df, random_sample))
mean(pull(df, gr_th_0))
mean(pull(df, num_char))
mean(pull(df, level))
```
The mean function cannot compute mean for variables num_char and level because they are non numerical variables (character, factor, etc.). The mean function computes the mean for variables random_sample and gr_th_0 because random_sample is a numerical variable and gr_th_0 is a logic vector, which could be interpreted as 0 and 1.

```{r apply as.numeric, eval = FALSE}
#turn character variable "num_char" into numerical variable
as.numeric(
        pull(df,num_char)
)

#turn factor vector "level" into numerical variable
as.numeric(
        pull(df,level)
)

#turn logical vector "gr_th_0" into numerical variable
as.numeric(
        pull(df,gr_th_0)
)
```
Applying the as.numeric function to the logical, factor, and character variables gives the elements in the original character vector (1,2,...8) in numerical form, corresponds "Low", "Medium", "High" to numbers 2,3,1, and corresponds 0,1 to False, True". It did this because as.numeric turns factor, logical, and character vectors into numeric vectors. This helps explain how the means couldn't be computed for character and factor vectors because they were non-numerical values. 

## Problem 2
```{r second code chunk, eval = TRUE, include = TRUE}
#convert character variable into numeric
num_char <- as.numeric(
            pull(df,num_char)
)

#mutiply the converted numerical variable with random sample
num_char * pull(df,random_sample)

#convert the logical vector to a factor
gr_th_0_factor <- as.factor(
                     pull(df,gr_th_0)
)

#mutiply result with random sample
gr_th_0_factor * pull(df,random_sample)

#convert logical vector to a factor and to numeric
gr_th_0_num <- as.numeric(
                      as.factor(
                          pull(df,gr_th_0)
))

#mutiply result with random sample
gr_th_0_num * pull(df,random_sample)
```

```{r, eval = TRUE, include = TRUE }
##load library tidyverse
library(tidyverse)

##set seed for random sample
set.seed(123)

##make a tibble dataframe with two random normal variables, a logical vector, and its numericalized and factorized form
df_1 <- tibble(x = rnorm(500, mean = 0, sd = 1), #sample following a normal distribution with mean = 0 and sd = 1
               y = rnorm(500, mean = 0, sd = 1), 
               #logical vector for sum greater than 1 
               gr_th_1 = x + y > 1, 
               #factorize above vector
               logic_num = as.numeric(gr_th_1), 
               #numericalize above vector
               logic_factor = as.factor(logic_num) 
               )
```
The dataset has `r nrow(df_1)` rows and `r ncol(df_1)` columns.
The mean of proportion of cases for which x + y > 1 is `r mean((pull(df_1,gr_th_1)) / (nrow(df_1)))` 

```{r make scatterplot, eval = TRUE, include = TRUE}
##create scatterplot of x and y color coded by logical vector
df_1 %>% 
  ggplot(aes(x = x, y = y, color = gr_th_1)) + 
  geom_point() ##dot for each datapoint + 
  theme(legend.position = "bottom") #legend position

##use ggsave to save plot
ggsave("plot_logical.png")

##create scatterplot of x and y color coded by "factorized" logical vector
plot_logic_factor <- 
  df_1 %>% 
  ggplot(aes(x = x, y = y, color = logic_factor)) + 
  geom_point() +
  theme(legend.position = "bottom") 
plot_logic_factor

ggsave("plot_logic_factor.png")

##create scatterplot of x and y color coded by "numericalized" logical vector
plot_logic_num <-
  df_1 %>% 
  ggplot(aes(x = x, y = y, color = logic_num)) + 
  geom_point(alpha = .5) +  
  theme(legend.position = "bottom")
plot_logic_num 

ggsave("plot_logic_num.png")
```
The scatterplots color-coded by the original logical vector, "factorized" logical vector, and "numericalized" logical vector are the same, because the variable is binary to begin with, and then numericalizing this logical vector just converts "TRUE" and "FALSE" into 1 and 0. Despite being presented on a continuous scale from 0 to 1 on the legend, the actual values are still just binary (0 and 1) as shown on the scatterplot with logic_num. 
Meanwhile, as.factor just turns this numerical logical vector (0,1 as numbers) back to factors, or levels of "0" and "1" (still binary, and thus only two colors for all datapoints in the scatterplot with logic factor).
